Effects of Antenna Beam Chromaticity on Redshifted 21~cm Power Spectrum and Implications for Hydrogen Epoch of Reionization Array

### 2016-03-11 ForegroundSimulations/ms.tex
Incorporate comments from Dave DeBoer
From 50bc153 (fgsim):
====================

In ms.tex,

(*) Incorporate comments from Dave DeBoer

When expressed in the coordinate system of power spectrum measurements described by the three-dimensional wavenumber ($k$), the foreground emission is restricted to a wedge-shaped region commonly referred to as the {\it foreground wedge} \citep{bow09,liu09,liu14a,liu14b,dat10,liu11,gho12,mor12,par12b,tro12,ved12,dil13,pob13,thy13,dil14}, whereas the EoR signal has spherical symmetry due to its isotropy which appears elongated along line of sight $k$ modes due to peculiar velocity effects when dominated by matter density perturbations during early stages of reionization.

@@

\citet{deb16} describe the HERA instrument in detail. Here, we provide a brief summary.

@@

In this study, we use the HERA-19 hexagonal array layout, a small subset of the planned layout. Without loss of generality, the analysis and results in this paper will apply to the full proposed layout as well.

@@

The delay spectrum has a close resemblance to cosmological H~{\sc i} spatial power spectrum and is defined as:
\begin{align}\label{eqn:delay-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{1}{\Omega\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \label{eqn:kperp-baseline}\\
  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, \label{eqn:kprll-delay}
\end{align}
where, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H~{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology, and following \citet{par14},
\begin{align}
  \Omega\Delta B &= \iint \left|A(\hat{\boldsymbol{s}},f)\,W(f)\right|^2\,\dif\Omega\,\dif f.
\end{align}

@@

The High Frequency Structural Simulator (HFSS) was used to model the dish and its angular response used in this study. The HFSS model used prime focus optics with a 14~m faceted parabola with a spar f/D ratio of 0.32.  The model has a 1~m central hole in the aluminum surface which is filled with a dielectric material similar to dry soil. The feed used a full PAPER dipole inside of a cylindrical backplane, the backplane is modeled as an aluminum surface. For the metal parts of the dipole, the discs were modeled as aluminum at the actual size, and the arms and terminals were modeled as copper. Dielectric stand-offs and supporting members were included. For the calculations, one pair of arms was excited using a modal port. These models cover a frequency range of 100--200~MHz in intervals of 1~MHz. Refer to \citet{deb16} for a complete description of the dish model.

@@

In a related series of papers, \citet{neb16} discuss the agreement of these simulated antenna beam patterns with actual measurements and \citet{ewa16} model the reflections and return loss expected in the proposed antenna-feed assembly. Our focus in this paper is to investigate chromaticity of antenna power patterns from the point of view of their impact on foreground contamination in delay power spectra.

@@

However, in all four cases considered here, the worst case signal-noise ratio lies above unity in most of this parameter space. It demonstrates that HERA will detect the EoR signal with a very high signal-noise ratio on all baselines in a vast majority of the pointings.

@@

Taking into account the effects of these instrumental systematics, if data is not limited by thermal noise, we demonstrate that HERA will detect EoR with high significance with s simple spectral weighting even under a conservative analysis criterion which does not involve {\it foreground removal} strategies. All baselines of HERA-19 will not only detect but also help in filtering highly likely EoR models.

>>>>

When expressed in the coordinate system of power spectrum measurements described by the three-dimensional wavenumber ($k$), the foreground emission is restricted to a wedge-shaped region commonly referred to as the {\it foreground wedge} \citep{bow09,liu09,liu14a,liu14b,dat10,liu11,gho12,mor12,par12b,tro12,ved12,dil13,pob13,thy13,dil14} bounded by the horizon limits \citep{par12b}, whereas the EoR signal has spherical symmetry due to its isotropy which appears elongated along line of sight $k$ modes due to peculiar velocity effects when dominated by matter density perturbations during early stages of reionization. The region of $k$-space excluding the {\it foreground wedge} is commonly referred to as the {\it EoR window} since it is expected to be relatively free of foreground contamination.

@@

\citet{deb16} describe the HERA instrument in detail. Here, we provide a brief summary of the telescope currently under construction on the South African SKA site.

@@

In this study, we use the HERA-19 hexagonal array layout, a small subset of the planned layout that are currently in use. Without loss of generality, the analysis and results in this paper will apply to the full proposed layout as well.

@@

The delay spectrum has a close resemblance to cosmological H~{\sc i} spatial power spectrum and is defined as:
\begin{align}\label{eqn:delay-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{1}{\Omega\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \label{eqn:kperp-baseline}\\
  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, \label{eqn:kprll-delay}
\end{align}
where, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $\boldsymbol{k}_\perp$ and $k_\parallel$ are the transverse (on the sky) and line-of-sight (into the sky) wavenumbers respectively, $f_{21}$ is the rest frame frequency of the 21~cm spin flip transition of H~{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight corresponding to $\Delta B$, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard terms in cosmology, and following \citet{par14},
\begin{align}
  \Omega\Delta B &= \iint \left|A(\hat{\boldsymbol{s}},f)\,W(f)\right|^2\,\dif\Omega\,\dif f,
\end{align}
which is related to the cosmic volume probed by the instrument.

@@

The High Frequency Structural Simulator (HFSS) was used to model the dish and its angular response used in this study. The HFSS model used prime focus optics with a 14~m faceted parabola with a spar f/D ratio of 0.32.  The model has a 1~m central hole in the aluminum surface which is filled with a dielectric material similar to dry soil. The feed used a full PAPER dipole inside of a cylindrical backplane, the backplane is modeled as an aluminum surface. For the metal parts of the dipole, the discs were modeled as aluminum at the actual size, and the arms and terminals were modeled as copper. Dielectric stand-offs and supporting members were included. For the calculations, one pair of arms was excited using a modal port. These models cover a frequency range of 90--210~MHz in intervals of 1~MHz \citep{deb16}.

@@

In a related series of papers, \citet{neb16} discuss the agreement of these simulated antenna beam patterns with actual measurements, \citet{ewa16} model the reflections and return loss expected in the proposed antenna-feed assembly, and \citet{pat16} present measurements quantifying the reflections. Our focus in this paper is to investigate chromaticity of antenna power patterns from the point of view of their impact on foreground contamination in delay power spectra.

@@

However, in all four cases considered here, the worst case signal-noise ratio lies above unity in most of this parameter space. It demonstrates that HERA should detect the EoR signal with a very high signal-noise ratio on all baselines in a vast majority of the pointings.

@@

Taking into account the effects of these instrumental systematics, if data is not limited by thermal noise, we demonstrate that HERA should detect EoR with high significance with s simple spectral weighting even under a conservative analysis criterion which does not involve {\it foreground removal} strategies. All baselines of HERA-19 will not only detect but also help in filtering highly likely EoR models.
